Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2014-12-19 doc/GMP2015/s03_drp.tex
Meera's change to problempinning. Small change to DRP

\begin{remark}
    \todo{this captures cluster minimality}
\end{remark}

>>>>

\begin{remark}
    \todo{Someone} \uncited discuss cluster minimality as a desirable property for children in a DR-plan. The canonical DR-plan in this paper intrinsically satisfies this property. If the children were not cluster minimal, then the union of children that form a larger cluster is itself isostatic and the original children were not vertex-maximal.
\end{remark}

### 2014-12-19 doc/GMP2015/s07_problempinning.tex
Meera's change to problempinning. Small change to DRP

We can use DR-plans to help in solving questions in designs for graphene and silica bi-layers \cite{silica_bilayers} \cite{sructure_of_2d_glass}. Specifically, we will look into the problem when you have a mono-layer and it is under-constrained. Our goal is to make such a layer isostatic. This can be done in a number of ways.

\begin{figure}\centering
    \includegraphics[width=0.4\linewidth]{img/Silica} \hspace{0.5cm}
    \raisebox{0.3\height}{\includegraphics[width=0.2\linewidth]{img/Silicon_tetrahedron}}
    \caption{Example of a single mono-layer of a Silica Oxygen glassy structure. This can be thought of as a multi-triangle-pin qusec where the Silicon atoms are the triangles and the Oxygen atoms are the pins. Not shown here are other layers stacked on this one. Each Silicon atom has the structure seen to the right, and so binds to another Oxygen atom in an adjacent mono-layer. Pictures taken from \cite{silica_figure} and \cite{tetra_silica_figure}.}
    \label{fig:silica_glass}
\end{figure}

\begin{enumerate}
    \item Pin together 2 under-constrained mono-layers in such a way that the resulting bi-layer becomes isostatic (see Figure \ref{fig:silica_glass}).
    \item Pin the boundary in such a way that the resulting system is isostatic.
    \item Build the structures by repeated subdivision to ensure they are isostatic at each layer (see Figure \ref{fig:subdivision})
\end{enumerate}

We will discuss item 2 in detail in this section. In each case, we are specifically interested in how to they can be applied to get a resulting isostatic structure with a small DR-plan, so a realization can be found quickly. To answer these questions, we first introduce the structures we model the problems with:

@@

    (b)
    \begin{subfigure}{0.8\linewidth}
        \includegraphics[width=\linewidth]{img/pentawesome}
    \end{subfigure}
    \caption{Examples of self-similarity in the form of repeated subdivision. In (a), a simple subdivision is done that does not guarantee isostatic. In (b), a more complicated scheme is used, ensuring that the resulting graph is both isostatic and non-tree-decomposable. Credit to \cite{subdivision_paper} for (a).}
    \label{fig:subdivision}
\end{figure}

\subsection{Body-Pin}

\begin{definition}
    A \dfn{body-hyperpin qusecs} is a qusecs where the object are rigid bodies that are pinned together by some number of pins.
\end{definition}

\begin{remark}
\label{rem:bodypin_is_barjoint}
    A body-hyperpin qusecs is a special case of bar-joint qusecs that we have been studying. As such, the DR-planning discussed in Section \ref{sec:DRP} is unchanged and the work in Section \ref{sec:recomb} will still go through with slight modifications.
\end{remark}

\begin{proof}
    We can replace each body that has only one pin by a single vertex. A body with 2 pins can be replaced by an edge. In general, a body with $n$ pins can be replaced by a 2-tree on $n$ vertices. When looking for a DR-plan, we treat each body as trivial, so they become the leaves of our plan. The optimal completion problem and approach of Section \ref{sec:DRP} are unchanged. The optimal parameterization problem in Section \ref{sec:DRP} now has an additional constraint that all edges in the 2-tree representation of the bodies must be removed together, not individually.
\end{proof}

For the rest of this section, we will only be dealing with the DR-plan of such qusecs. As such, we will refer only to the underlying graph of the qusecs. We now go over 2 sub-classes of body-hyperpin graphs for modeling Examples 4 and 5 in Section \ref{sec:intro}.

\begin{definition}
\label{def:body-pin}
    A body-pin graph is a body-hyperpin graph with the following conditions:
    \begin{itemize}
        \item Each pin is shared by at most 2 bodies
        \item No 2 bodies share more than one pin
    \end{itemize}
    Such a body-pin graph can also be seen as a \dfn{body-bar graph} where the bodies are the original bodies and each pin represents 2 bars from one body to another. Such body-bar graphs with 1 and 2-dof can be characterized by being $(3,4)$ and $(3,5)$-tight respectively \cite{Lee:2007:PGA} \cite{streinu2009sparse}.
\end{definition}

\begin{definition}
    A \dfn{multi-triangle-pin graph} is a body-hyperpin graph where each body is a triangle, meaning it is pinned in at most 3 places. This is also represented as a hyper-graph where each pin is a vertex and each triangle represents a tri-hyperedge. For such hyper-graphs, 1 and 2-dof can be characterized by $(2,4)$ and $(2,5)$-tightness respectively \cite{Lee:2007:PGA} \cite{streinu2009sparse}.
\end{definition}

Body-pin graphs are of particular interest to us in the context of Example 4 in Section \ref{sec:intro}. Multi-triangle-pin graphs can be used to represent the silica bi-layers and glassy structures described in Example 5 of Section \ref{sec:intro}, where each triangle is the junction of ``disks'' in the plane (see Figure \ref{fig:silica_glass}). Typically, these systems are not isostatic, so to relate the work of this paper to the systems, we define a slightly different kind of DR-plan:

@@

We can develop the notion of a canonical DR-plan for certain $(k,l)$-tightness conditions similar to the way we did in Section \ref{sec:DRP}.

\begin{observation}
\label{obs:bodypin_drp}
    For the 1-dof body-hyperpin graphs described above, a DR-plan node will follow one of the following:

    \begin{itemize}
        \item Its children will be 2 proper vertex maximal 1-dof graphs that intersect on another 1-dof graph
        \item Its children will be a number of proper maximal 1-dof sub-graphs, joined pairwise by at most one pin
    \end{itemize}
\end{observation}

@@

\begin{remark}\label{rem:1dofcanon}
    All $(k,l)$-tight canonical DR-plans are optimal. We can find such a DR-plan in the same time complexity as the $(2,3)$-tight case for bar and joint graphs discussed in Section \ref{sec:DRP}. This is due to the fact that such $(k,l)$-tight graphs are matroidal and have a pebble game, shown in \cite{Lee:2007:PGA}.
\end{remark}

\begin{figure*}\centering
\begin{subfigure}{0.2\linewidth}\centering
    \includegraphics[width=\linewidth]{img/bodypin}
    \caption{}
\end{subfigure}
%
\begin{subfigure}{0.2\linewidth}\centering
    \input{img/tikz/bodypin_graph.tikz}
    \caption{}
\end{subfigure}
%
\begin{subfigure}{0.35\linewidth}\centering\scriptsize
    \input{img/tikz/bodypin_drp.tikz}
    \caption{}
\end{subfigure}
%
\begin{subfigure}{0.2\linewidth}\centering
    \includegraphics[width=\linewidth]{img/bodypin2}
    \caption{}
\end{subfigure}
%
\caption{(a) is a 1-dof body-pin graph. (b) is the corresponding body-bar graph explained in Definition \ref{def:body-pin}. (c) is the 1-dof DR-plan for the graph. In this case, to obtain an isostatic system, we would need to add a body and 2 pins to one of the nodes in the second level. (d) is the result of adding one such body to the bold-faced node.}
%
\end{figure*}

\begin{theorem}
\label{thm:1dofcase}
    Given a 1-dof body-pin or multi-triangle graph $G$ and corresponding 1-dof DR-plan $T$, there is a quadratic algorithm for the 1-dof optimal completion problem of Section \ref{sec:recomb} for body-pin and multi-triangle-pin graphs.
\end{theorem}

@@

\begin{observation}
\label{obs:algebraic_completion}
    While the proof for Theorem \ref{thm:1dofcase} gives us a DR-plan with minimum fanin, we did not consider the algebraic complexity of realizing the DR-plan. If that is preferred, we can no longer only consider the deepest nodes.
\end{observation}

>>>>

We can use qusecs DR-plans to design materials such as disordered
graphene and silica bi-layers \cite{silica_bilayers}
\cite{sructure_of_2d_glass}. We will investigate a more specific problem in
a somewhat more general setting:
the problem of finding boundary conditions (additional constraints) to
add to an underconstrained monolayer to make it
isostatic.
This can be done in a number of ways.

\begin{figure}\centering
    \includegraphics[width=0.4\linewidth]{img/Silica} \hspace{0.5cm}
    \raisebox{0.3\height}{\includegraphics[width=0.2\linewidth]{img/Silicon_tetrahedron}}
    \caption{Example of a single monolayer of a Silica (silicon and oxygen)
    glassy structure. This can be thought of as a triangle-multipin qusec
where the Silicon atoms are the triangles and the Oxygen atoms are the pins.
Not shown here are other monolayers stacked on this one. Each Silicon atom
has the structure seen to the right, and so binds to another Oxygen atom in
an adjacent monolayer. Pictures taken from \cite{silica_figure} and \cite{tetra_silica_figure}.}
    \label{fig:silica_glass}
\end{figure}

\begin{enumerate}
    \item Pin together 2 under-constrained monolayers in such a way that the resulting bi-layer becomes isostatic (see Figure \ref{fig:silica_glass}).
    \item Pin the boundary of (or in general, add constraints to) a layer
        (possibly a genus 0 monolayer) so that it becomes isostatic
    \item Design a broader class of structures to ensure they are isostatic,
        self-similar (via some subdivision rule) and in addition isostatic at each
        level of the subdivision (see Figure \ref{fig:subdivision})
\end{enumerate}

In all cases,
we are specifically interested in how to add additional constraints
resulting isostatic structure with a small DR-plan, so a realization can be
found and stress and other property design can be efficient.
To answer these questions, we first introduce the
qusecs that are used to model these materials.
We will discuss Item 2 in detail in this section.

@@

    (b)
    \begin{subfigure}{0.8\linewidth}
        \includegraphics[width=\linewidth]{img/pentawesome}
    \end{subfigure}
    \caption{Examples of self-similarity via repeated subdivision. In (a),
    simple subdivision that does not guarantee isostaticity. In (b), a more
complicated scheme is used, ensuring that the resulting graph is isostatic
but not  tree-decomposable. Credit to \cite{subdivision_paper} for (a).}
    \label{fig:subdivision}
\end{figure}

\subsection{Body-Hyperpin qusecs}

\begin{definition}
    A \dfn{body-hyperpin qusecs} is a constraint system where the objects are rigid bodies
    some subsets of which are pinned together by pins.
\end{definition}

\begin{remark}
\label{rem:bodypin_is_barjoint}
    A body-hyperpin qusecs is a special case of bar-joint qusecs of the
    first 5 sections. As such, the DR-planning for isostatics discussed in Section
    \ref{sec:DRP} is unchanged and the results of Section \ref{sec:recomb} will still go through with slight modifications.
\end{remark}

\begin{proof}
    We can replace each body that has only one pin by a single vertex. A
    body with 2 pins can be replaced by an edge. In general, a body with $n$
    pins can be replaced by a 2-tree on $n$ vertices. When finding a
    DR-plan, we treat each body as trivial, so they become the leaves of our
    plan. The optimal recombination problem and approach of Section
    \ref{sec:recomb} are unchanged. The optimal completion via the optimal modification problem in
Section \ref{sec:recomb} now has an additional constraint that all edges in the 2-tree representation of the bodies must be removed together, not individually.
\end{proof}

For the rest of this section, we will only be dealing with the DR-plan of
such qusecs. As such, we will refer only to the underlying hypergraph of the qusecs. We now
introduce 2 sub-classes of body-hyperpin graphs for modeling Examples 4 and
5 in Section \ref{sec:intro}, for which the optimal completion problem is
significantly easier.

\begin{definition}
\label{def:body-pin}
    A body-pin graph is a body-hyperpin graph with the following conditions:
    \begin{itemize}
        \item Each pin is shared by at most 2 bodies
        \item No 2 bodies share more than one pin
    \end{itemize}
    Such a body-pin graph can also be seen as a \dfn{body-bar graph} where
    the bodies are the original bodies and each pin represents 2 bars from
    one body to another. Such body-bar graphs with 1 and 2-dof can be
    characterized by being $(3,4)$ and $(3,5)$-tight respectively
    \cite{Lee:2007:PGA} \cite{streinu2009sparse},
defined in Section  \ref{sec:preliminaries.}
\end{definition}

\begin{definition}
    A \dfn{triangle-hyperpin graph} is a body-hyperpin graph where each body
    is a triangle, i.e., it shares at most 3pins with other bodies. This is also represented as a hyper-graph where each pin is a vertex and each triangle represents a tri-hyperedge. For such hypergraphs, 1 and 2-dof can be characterized by $(2,4)$ and $(2,5)$-tightness respectively \cite{Lee:2007:PGA} \cite{streinu2009sparse}.
\end{definition}

Body-pin graphs are of particular interest to us in the context of Example 4
in Section \ref{sec:intro}. Triangle-multipin graphs can be used to
represent the silica bi-layers and glassy structures described in Example 5
of Section \ref{sec:intro}, where each triangle is the junction of ``disks''
in the plane (see Figure \ref{fig:silica_glass}). Typically, these systems
are not isostatic, so to relate the work of this paper to the systems, we
define a slightly different kind of DR-plan using the notion of
$(k,l)$-sparsity and tightness.

@@

We can develop the notion of a canonical DR-plan when the hypergraph is
$(k,l)$-sparse (i.e., combinatorially independent) using certain
notion of trivial and non-trivial intersections and
$(k,l)$-tightness conditions similar to the way we did in Section
\ref{sec:DRP}, provided
    such $(k,l)$-sparse graphs are matroidal -- conditions given in \cite{Lee:2007:PGA}.
    In particular, we can define canonical DR-plans with similar properties
    for the 1 and 2 dof body-pin and triangle-hyperpin systems defined
    above.

\begin{observation}
\label{obs:bodypin_drp}
    For the 1-dof body-pin graphs described above, a $(3,4)$-sparsity
    corresponds to canonical DR-plan node
    satisfies one of the following:

    \begin{itemize}
        \item Its children  are 2 proper vertex maximal 1-dof graphs that intersect on
            another 1-dof graph
        \item Its children are a number of proper maximal 1-dof sub-graphs,
            sharing at most one body
    \end{itemize}
\end{observation}

@@

The strong Church-Rosser property holds making all canonical DR-plans
optimal:
%
\begin{remark}\label{rem:1dofcanon}
    When the input is independent,
    all $(3,4)$-tight canonical DR-plans are optimal. We can find such a
    DR-plan in the same time complexity as the $(2,3)$-tight case for bar
    and joint graphs discussed in Section \ref{sec:DRP}.
    This is because
    such $(3,4)$-tight graphs are matroidal and have a pebble game, shown in \cite{Lee:2007:PGA}.
\end{remark}

\begin{figure*}\centering
\begin{subfigure}{0.2\linewidth}\centering
    \includegraphics[width=\linewidth]{img/bodypin}
    \caption{}
\end{subfigure}
%
\begin{subfigure}{0.2\linewidth}\centering
    \input{img/tikz/bodypin_graph.tikz}
    \caption{}
\end{subfigure}
%
\begin{subfigure}{0.35\linewidth}\centering\scriptsize
    \input{img/tikz/bodypin_drp.tikz}
    \caption{}
\end{subfigure}
%
\begin{subfigure}{0.2\linewidth}\centering
    \includegraphics[width=\linewidth]{img/bodypin2}
    \caption{}
\end{subfigure}
%
\caption{(a) is a 1-dof body-pin graph. (b) is the corresponding body-bar graph explained in Definition \ref{def:body-pin}. (c) is the 1-dof DR-plan for the graph. In this case, to obtain an isostatic system, we would need to add a body and 2 pins to one of the nodes in the second level. (d) is the result of adding one such body to the bold-faced node.}
%
\end{figure*}
The above discussion leads to the main theorem:
\begin{theorem}
\label{thm:1dofcase}
    Given a 1-dof body-pin or triangle-multipin graph $G$ and corresponding
    1-dof DR-plan $T$, there is a quadratic algorithm for the 1-dof optimal
    completion problem of Section \ref{sec:recomb}.
\end{theorem}

@@

\begin{observation}
\label{obs:algebraic_completion}
    While the proof for Theorem \ref{thm:1dofcase} gives us a DR-plan for
    the isostatic completion with minimum fanin, (a reasonable
    measure of algebraic complexity), a more nuanced
    measure that treats solutions of 1-dof and 2-dof systems as
    1- or 2- parameter families would no longer permit the algorithm given
    in the proof. In particular the complexity of the standard algorithm in the k-dof
    case would be exponential in $k$ (even if that case were matroidal and
    an optimal DR-plan is known).
\end{observation}
